---
title: R in Space - Basic mapping
---


# Basic mapping


## Plot methods for `sp` objects

Let's first get some data:

```{r getdatabel, message = FALSE}
library(sp)
library(sf)
library(raster)
bel1 <- getData("GADM", country = "BEL", level = 1, path =  "data")
bel2 <- getData("GADM", country = "BEL", level = 2, path =  "data")
```

Quick plot:

```{r plotsp1}
plot(bel1)
```

Two quick plots:

```{r plotsp2}
par(mfrow = c(1, 2), mar = c(0, 0, 0, 0))
plot(bel1, col = 1:3)
plot(bel2, lty = 2)
```

Slightly customized plots:

```{r plotsp3}
plot(bel1, col = "grey85", main = "Belgium")
plot(bel2, lwd = .5, add = TRUE)
points(coordinates(bel2), pch = 19, cex = .4)
city <- coordinates(bel2)[1, ]
text(city[1], city[2], labels = "Bruxelles", pos = 3, col = "#9728eb")
```




## plot methods for `sf` objects

Let's first get two `sf` object:

```{r map_sf0, message = FALSE}
bel1_sf <- st_as_sf(bel1)
bel2_sf <- st_as_sf(bel2)
```

Let's try a simple plot.

```{r map_sf1, message = FALSE}
plot(bel1_sf)
```

It actually plots one map per variable and colors the polygon according
to the values. To obtain something similar to what is found in `sp`,
an `sfc` object must be used:


```{r map_sf2, message = FALSE}
plot(st_geometry(bel1_sf), lwd = 1.5, col = "grey85")
# same as :
# plot(bel1_sf$geometry, lwd = 1.5, col = "grey85")
plot(st_geometry(bel2_sf), lwd = .5, add = TRUE)
```


## plot methods for `raster` objects

Let's get a raster layer of altitude in Belgium using again `getData()`.

```{r getdata_alt, message = FALSE}
alt <- getData("alt", country = "Belgium", path =  "data")
```

```{r plot_alt, message = FALSE}
plot(alt)
```


We can also add different layers of information on this maps:



```{r plot_alt2, message = FALSE}
plot(alt)
contour(alt, add = TRUE)
```

Or reclassify and change the color palette:


```{r plot_alt3, message = FALSE}
# Assign color to a object
myCol <- terrain.colors(3)
plot(alt,
     breaks = c(0, 200, 400, 700),
     col = myCol,
     main = "Altitude map of Belgium")
```

## plotting together `sf` and `raster` objects

```{r plot_alt_sf, message = FALSE}
# Create a new color palette
library(RColorBrewer)
myPal <- brewer.pal(n = 9, name = "Greens")

# Plot the altitude raster
plot(alt,
  col = myPal,
  xlab = "Longitude",
  ylab = "Latitude")

# Add Belgium boundary map
plot(st_geometry(bel2_sf), add = TRUE)

# Add a point at the centroid of Bruxelles
brux <- bel2_sf[bel2_sf$NAME_1 == "Bruxelles",]

plot(st_geometry(st_centroid(brux)), pch = 20, add = TRUE)

# Add a label for Bruxelles on the map
text(st_coordinates(st_centroid(brux)), labels = "Bruxelles", pos = 3, col = "#9728eb")
```


## Package `mapview`

```{r mapview, message = FALSE, fig.width = 9.5}
library(mapview)
mapview(bel1)@map
```

